# Database Creation

Data collected for this project was: collection sites, release sites, length, girth, PIT tag ID, and genetic clips. below is a flow chart of tables created in Rstudio and SQLite.

![](images/2025-01-30%20(3).png)

## Create Lamprey Table

Code used in SQLite to create the Lamprey Table. Table includes Lamprey ID(primary key), Capture Sites(Foreign Key), Release Sites, Genetic, and PIT Tag \#(Foreign Key).

```{sql create lamprey table, eval=FALSE, echo = TRUE}
CREATE TABLE Lamprey (
    [Lamprey ID]   TEXT    PRIMARY KEY
                           NOT NULL,
    [Capture Site] TEXT    NOT NULL
                           REFERENCES Captures ([Capture Site]),
    [Release Site] TEXT    NOT NULL,
    Genetic        INTEGER,
    [PIT Tags #]   TEXT    REFERENCES Lamprey ([PIT Tags #]) 
);
```

I copied and pasted the raw data from my excel file into the Lamprey Table.

## Create Captures Table

Code used in SQLite to create Capture Table. Table includes: Capture ID(Primary Key), Lamprey ID (Foreign Key), Capture Site(Foreign Key), Date.

```{sql create capture table, echo = TRUE, eval = FALSE}
CREATE TABLE Captures (
    [Capture ID]   TEXT CONSTRAINT [duplicated values] PRIMARY KEY,
    [Lamprey ID]   TEXT REFERENCES Lamprey ([Lamprey ID]) MATCH SIMPLE,
    [Capture Site] TEXT REFERENCES Capture_Site ([Capture Site]) MATCH SIMPLE,
    Date           TEXT
);
```

## Create Capture Site Table

Code used in SQLite to create Capture Site Table. Table includes: Capture Site(Primary Key), Capture UTM-X, Capture UTM-Y.

```{sql create capture site table, echo = TRUE, eval = FALSE}
CREATE TABLE Capture_Site (
    [Capture Site]  TEXT    NOT NULL
                            PRIMARY KEY
                            UNIQUE ON CONFLICT REPLACE,
    [Capture UTM-X] NUMERIC NOT NULL,
    [Capture UTM-Y] NUMERIC NOT NULL
);
```

## Create Tag Table

Code used in SQLite to create Tag Table. Table includes: PIT Tag \#(Primary Key), Status, Lamprey ID (Foreign Key).

```{sql create tag table, echo = TRUE, eval = FALSE}
CREATE TABLE Tags (
    [PIT Tag #]  TEXT PRIMARY KEY
                      NOT NULL,
    Status       TEXT,
    [Lamprey ID] TEXT REFERENCES Lamprey ([Lamprey ID]) MATCH SIMPLE
);
```

## Create Morphometrics Table

Code used in SQLite to create Morphometrics Table. Table includes: Measurement ID(Primary Key), Length, Girth, Genetic, Lamprey ID(Foreign Key).

```{sql create morphometrics table, echo = TRUE, eval = FALSE}
CREATE TABLE Morphometrics (
    [Measurement ID] TEXT    PRIMARY KEY
                             NOT NULL,
    [Length (mm)]    NUMERIC,
    [Girth (mm)]     NUMERIC,
    Genetic          INTEGER,
    [Lamprey ID]     TEXT    REFERENCES Lamprey ([Lamprey ID]) MATCH SIMPLE
                             NOT NULL
);
```

After creating the tables I copied the information into the tables from my raw excel file of my data.

## Connecting SQLite to Rstudio to upload my files

Load packages.

```{r connnecting SQLite to Rstudio, echo=TRUE, eval = TRUE}
library(DBI)
library(RSQLite)
library(bookdown)
```

Connected Rstudio to my SQLite File

```{r Connecting RStudio to SQLite, eval = TRUE, echo = TRUE}
Final_db <- dbConnect (RSQLite::SQLite(), "Final.db")
```

## Uploaded tables into Rstudio

Importing tables from SQLite to Rstudio.

```{r}
Morphometrics <- dbGetQuery (conn = Final_db,
            statement = "SELECT * FROM Morphometrics;")

Capture_Site <- dbGetQuery (conn = Final_db,
            statement = "SELECT * FROM Capture_Site;")
Captures <- dbGetQuery (conn = Final_db,
            statement = "SELECT * FROM Captures;")
Lamprey <- dbGetQuery (conn = Final_db,
            statement = "SELECT * FROM Lamprey;")
Tags <- dbGetQuery (conn = Final_db,
            statement = "SELECT * FROM Tags;")
```

Printing tables to double check uploads.

```{r}
head(Morphometrics)
tail(Morphometrics)
```

```{r}
print(Capture_Site)
```

```{r}
head(Captures)
tail(Captures)
```

```{r}
head(Lamprey)
tail(Lamprey)
```

```{r}
head(Tags)
tail(Tags)
```
