# Data Cleaning {#cross}

### Data prepping for:

1.  Average length per site

    1.  Joined the tables "Morphometrics" and "Lamprey"

    2.  removed columns not needed and reorganized where Capture site column came before Length.

    3.  renamed "Capture Site" and "Length (mm)" to "capture_site" and "length" to remove spaces

    4.  grouped the columns by capture site

    5.  calculated average length of lamprey per captured site

        ```{r Create table of average length per site, eval = FALSE, echo=TRUE}
        #Creating table with average length per site
        average_length_per_site <- left_join(Morphometrics, Lamprey, by = "Lamprey ID") %>% 
          select(-"Measurement ID",-"Genetic.x", -"Release Site", -"Genetic.y", -"Release Site", -"PIT Tags #",
                 -"Lamprey ID", -"Girth (mm)") %>% 
          select("Capture Site", "Length (mm)") %>% 
         rename(capture_site = "Capture Site", length = "Length (mm)") %>% 
          group_by(capture_site) %>% 
          summarize(average = mean(length))
        ```

2.  average girth per capture site

    1.  Joined the tables "Morphometrics" and "Lamprey"

    2.  removed columns not needed and reorganized where Capture site column came before girth.

    3.  renamed "Capture Site" and "Girth (mm)" to "capture_site" and "girth" to remove spaces

    4.  grouped the columns by capture site

    5.  calculated average length of lamprey per captured site

        ```{r creating table for average girth per site, eval = FALSE, echo = TRUE}
        #create table with average girth per site
        average_girth_per_site <- left_join(Morphometrics, Lamprey, by = "Lamprey ID") %>% 
          select(-"Measurement ID",-"Genetic.x", -"Release Site", -"Genetic.y", -"Release Site", -"PIT Tags #",
                 -"Lamprey ID", -"Length (mm)") %>% 
          select("Capture Site", "Girth (mm)") %>% 
          rename(capture_site = "Capture Site", girth = "Girth (mm)") %>% 
          group_by(capture_site) %>% 
          summarize(average = mean(girth))
        ```

3.  average length per release site

    1.  Joined the tables "Morphometrics" and "Lamprey"

    2.  removed columns not needed and reorganized where release site column came before Length.

    3.  renamed "Release Site" and "Length (mm)" to "release_site" and "length" to remove spaces

    4.  Reorganized columns with "release_site" being first before "length"

    5.  renamed "Beaver" responses to "BEAV" to match the data in the table

    6.  grouped the columns by release site

    7.  calculated average length of lamprey per captured site

        ```{r creating table for average length at release site, echo = TRUE, eval = FALSE}
        #create table with average length per release site
        average_length_per_release_site <- left_join(Morphometrics, Lamprey, by = "Lamprey ID") %>% 
          select(-"Measurement ID",-"Genetic.x", -"Genetic.y", -"Capture Site", -"PIT Tags #",
                 -"Lamprey ID", -"Girth (mm)") %>% 
          select("Release Site", "Length (mm)") %>% 
          rename(release_site = "Release Site", length = "Length (mm)") %>% 
          mutate(release_site = case_when(
            release_site == "Beaver" ~ "BEAV",TRUE ~ release_site)) %>% 
          group_by(release_site) %>% 
          summarize(average = mean(length))
        ```

4.  average girth per release site

    1.  Joined the tables "Morphometrics" and "Lamprey"

    2.  removed columns not needed and reorganized where release site column came before girth.

    3.  renamed "Release Site" and "Girth (mm)" to "release_site" and "girth" to remove spaces

    4.  Reorganized columns with "release_site" being first before "girth"

    5.  renamed "Beaver" responses to "BEAV" to match the data in the table

    6.  grouped the columns by release site

    7.  calculated average length of lamprey per captured site

        ```{r create table for average girth per release site, echo = TRUE, eval = FALSE}
        #create table with average girth per relase site
        average_girth_per_release_site <- left_join(Morphometrics, Lamprey, by = "Lamprey ID") %>% 
          select(-"Measurement ID",-"Genetic.x", -"Genetic.y", -"Capture Site", -"PIT Tags #",
                 -"Lamprey ID", -"Length (mm)") %>% 
          select("Release Site", "Girth (mm)") %>% 
          rename(release_site = "Release Site", girth = "Girth (mm)") %>% 
          mutate(release_site = case_when(
            release_site == "Beaver" ~ "BEAV",TRUE ~ release_site)) %>% 
          group_by(release_site) %>% 
          summarize(average = mean(girth))
        ```

5.  number of captured lamprey per day at Sherears Falls

    1.  removed the columns that were not need in the "Captures" table

    2.  renamed "Capture Site" to "capture_site" to remove the spaces

    3.  changed the "Shears" outputs to "Sherears" for correct spelling errors.

    4.  filtered for Sherears captured lamprey only

    5.  grouped the lamprey by the day they were captured

    6.  tallied by how many were captured per day

        ```{r how many lamprey captured per day, echo = TRUE, eval = FALSE}
        # how many lamprey captured per day
        capture_per_day_sherears <- Captures %>% 
          select(-"Capture ID", -"Lamprey ID") %>% 
          rename(capture_site = "Capture Site") %>% 
          mutate(capture_site = case_when(
            capture_site == "Shears" ~ "Sherears", TRUE ~ capture_site)) %>% 
          filter(capture_site == "Sherears") %>% 
          group_by(Date) %>% 
          tally
        ```

6.  Suvival of tagging at the Warm Springs Hatchery Site

    1.  Joined the "Lamprey" and "Tags" table

    2.  removed all the unwanted columns

    3.  renamed "Release Site" to "release_site" to remove spaces

    4.  added column "survival" into the table and set alive = 1 and dead = 0

    5.  filtered out all sites besides "WS Hatch"

    6.  calculated survival by summing Alive Lamprey by number of released at "WS Hatch) and multiplying by 100 to get an percentage

        ```{r survival after tagging, echo = TRUE, eval = FALSE}
        #survival after tagging
        tag_survival_capture_site <- full_join(Lamprey, Tags) %>% 
          select(-"Lamprey ID", -"Capture Site", -"Genetic", -"PIT Tag #",-"PIT Tags #") %>% 
          rename(release_site = "Release Site") %>% 
          mutate(survival = ifelse(Status == "Alive", 1, 0)) %>% 
          filter(release_site == "WS Hatch") %>% 
          select(-"Status") %>% 
          summarize((sum(survival)/40))*100
        ```
